\thetitle

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
last minor changes before overhauling reconstruction error bounds

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\usepackage{mathtools} % for \coloneqq

@@

\begin{abstract}
  We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
  We conjecture that $r_d=2d-1$, which we have verified for all $d\le100$ using numerical experiments.
  We then provide an algorithm with $O\p{r^3}$ runtime for certifying strict upper bounds on the error with which measurements of spin projection along $r$ measurement axes can be used to reconstruct an unknown quantum state.
  Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error [\red{note: this part is not yet complete}].
  [\red{todo: maybe look into a simplified tomography procedure for pure states?]}
\end{abstract}

@@

Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
Since
\begin{enumerate*}
\item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
\item measuring spin projections along any axis simultaneously provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
\end{enumerate*}
one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any degree-$\ell$ function with $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ would allow recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.  However, this condition is stronger than that necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes.}.
Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
Nonetheless, in the following section we show by example that $2d-1$ measurement axes suffice for all $d\le100$. [\red{modify sentence as appropriate after writing the next section}]

@@

A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\v\Pi_V=\set{\Pi_j^V}$ that will be measured in an experiment, where for shorthand we use a combined index $j\simeq\p{\v v_j,\mu_j}$ that specifies both a measurement axis $\v v_j$ and an outcome $\mu_j$, and for clarify we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
The projectors $\v\Pi_V$ can be used to construct the overlap (Gram) matrix
\begin{align}
  G_V = \sum_{i,j=1}^{\abs{V}\times d} \obk{\Pi_i^V|\Pi_j^V} \op{i}{j}.
  \label{eq:overlap_matrix}
\end{align}
Eigenvectors of $G_V$ define linearly independent linear combinations of the projectors in $\v\Pi_V$.
If the projectors in $\v\Pi_V$ span the entire space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
Indexing these eigenvalues $G^V_k$ and the corresponding (normalized) eigenvectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$,
we can construct the orthonormal operators
\begin{align}
  Q^V_k \equiv \sum_j g^V_{kj} \Pi_j^V,
  &&
  g^V_{kj} \equiv \f{x^V_{kj}}{\sqrt{G^V_k}}.
\end{align}
These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
\begin{align}
  \rho = \sum_{k=1}^{d^2} \rho_k^V Q_k^V,
  &&
  \rho_k^V \equiv \bk{Q^V_k}_\rho.
\end{align}
Given empirical estimates $E_j^V$ of the expectation values $\bk{\Pi_j^V}_\rho$, an empirical estimate $\tilde\rho$ of $\rho$ is then
\begin{align}
  \tilde\rho \equiv \sum_k \tilde\rho^V_k Q^V_k,
  \label{eq:reconstructed_state}
\end{align}
where
\begin{align}
  \tilde\rho^V_k \equiv \sum_j g^V_{kj} E_j^V
  \approx \sum_j g^V_{kj} \bk{\Pi_j^V}_\rho
  = \bk{Q^V_k}_\rho
  = \rho^V_k.
\end{align}
The eigenvalues and eigenvectors of the overlap matrix $G_V$ provide strict upper bounds on the distance between the empirical estimate $\tilde\rho$ and the true state $\rho$.
If the estimates $E^V_j$ approximate $\bk{\Pi^V_j}_\rho$ to within an error $\epsilon$, i.e.
\begin{align}
  \abs{E^V_j-\bk{\Pi^V_j}_\rho} \le \epsilon,
\end{align}
then the error in the estimate $\tilde\rho^V_k$ of $\rho^V_k$ is
\begin{align}
  \abs{\tilde\rho^V_k-\rho^V_k}
  &= \abs{\sum_j g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}} \\
  &\le \sum_j \abs{g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}} \\
  &\le \epsilon \norm{\v g^V_k}_1,
\end{align}
where $\norm{\v g}_1=\sum_j\abs{g_j}$ is the 1-norm of the vector $\v g$.
Using the fact that the operators $Q^V_k$ are orthonormal, we can bound the 2-norm (Euclidean) distance $\norm{\tilde\rho-\rho}_2$ between $\tilde\rho$ and $\rho$ as
\begin{align}
  \norm{\tilde\rho-\rho}_2^2
  = \sum_k \abs{\tilde\rho^V_k - \rho^V_k}^2
  \le \epsilon^2 \N_V^2,
  \label{eq:bound}
\end{align}
where
\begin{align}
  \N_V^2 \equiv \min_U \sum_k \norm{U\v g^V_k}_1^2
  = \min_U \sum_k \f{\norm{U\v x^V_k}_1^2}{G^V_k}
  \label{eq:norm}
\end{align}
is a norm uniquely determined by the set of measurement axes $V$, with minimization performed over unitary transformations $U$ of all $\v g^V_k$ and $\v x^V_k$ to account for the fact that their 1-norms are generally basis-dependent.
Seen otherwise, this minimization accounts for the fact that we can repeat the analysis between Eqs.~\eqref{eq:overlap_matrix} and \eqref{eq:bound} with any $G_V\to U G_V U^\dag$, where $U$ is an arbitrary unitary that mixes the projectors in $\v\Pi_V$.

To better understand the bound in Eq.~\eqref{eq:bound}, we observe that by a change of operator basis, or equivalently a particular choice of unitary $U_T$, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
Orthogonality of transition operators with different degrees $\ell$ then implies that the overlap matrix
\begin{align}
  G_V^{\t{trn}} \equiv U_T G_V U_T^\dag
  = \sum_{\ell,\v v,\v v'} \obk{T_{\v v\ell,0}|T_{\v v'\ell,0}}
  \op{\v v\ell}{\v v'\ell}
\end{align}
has a block-diagonal structure, with $d$ blocks of size $\abs{V}\times\abs{V}$.
Eigenvectors of $G_V$ can therefore have at most $\abs{V}$ nonzero components in the transition operator basis, which implies that
\begin{align}
  \N_V \le \sum_k \f{\norm{U_T\v x^V_k}_1^2}{G^V_k}
  \le \sum_k \f{\abs{V}\,\norm{U_T\v x^V_k}_2^2}{G^V_k}
  = \sum_k \f{\abs{V}}{G^V_k}.
\end{align}
Note that these sums are performed over the $d^2$ largest eigenvalues of $G_V$ (or $G_V^{\t{trn}}$), and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
Altogether, diagonalizing each of $d$ blocks of $G_V^{\t{trn}}$ takes a time that is $O(\abs{V}^3)$, so an upper bound for $\N_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
[\red{Note: rather than an absolute error $\epsilon$ on experimental measurements, we can assume that $E_j$ is equal to $\bk{\Pi_j}_\rho$ up to Gaussian noise with variance $\epsilon^2$.  In that case, rather than a strict error bound we can make statements about the expected (squared) distance between $\tilde\rho$ and $\rho$, which turns out to be simply $\epsilon^2\sum_k1/G^V_k$.  In retrospect, I think that this is a much nicer result, but I haven't written it up yet, which will require me to restructure this section somewhat.}]

The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the cost function
\begin{align}
  \tilde\N_V^2 \equiv \sum_k \f{\norm{\v y^V_k}_1^2}{G^V_k}
  \ge \N_V^2,
\end{align}
where $\v y^V_k$ are the (normalized) eigenvectors of the overlap matrix $G_V^{\t{trn}}$ in the transition operator basis, with corresponding eigenvalues $G^V_k$.
If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\tilde\N_V^2$.
Upon choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi^V_j}_\rho$.
If these measurements can approximate the expectation values $\bk{\Pi^V_j}_\rho$ to within a known error $\epsilon$, then the empirical estimate $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance of $\epsilon\N_V$.
If the empirical estimate $\tilde\rho$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced via maximum-likelihood corrections \cite{smolin2012efficient}.
In terms of classical computing costs before and after the estimation of expectation values $\bk{\Pi^V_j}_\rho$, this tomography protocol is bottlenecked by the need to diagonalize the overlap matrix $G_V^{\t{trn}}$, which has $d$ blocks of size $\abs{V}\times\abs{V}$.
Altogether, this protocol has an $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.

@@

As a last point, we mention that the cost function $\tilde\N_V^2$ can also evaluate the quality of measurement axes chosen using existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-cost measurement axes with higher probability.
Two observations may explain the poor performance of the sparse signal recovery methods: first, that spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely with a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of sparse signals.
Second, spin qudit tomography essentially requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

>>>>

@@

\begin{abstract}
  We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
  We conjecture that $r_d=2d-1$, which we have verified for all $d\le120$ using numerical experiments.
  We then provide an algorithm with $O\p{r^3}$ runtime for certifying strict upper bounds on the error with which measurements of spin projection along $r$ measurement axes can be used to reconstruct an unknown quantum state.
  Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error [\red{note: this part is not yet complete}].
  [\red{todo: maybe look into a simplified tomography procedure for pure states?]}
\end{abstract}

@@

Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
Since
\begin{enumerate*}
\item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
\item measuring spin projections along any axis simultaneously provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
\end{enumerate*}
one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any degree-$\ell$ function with $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ would allow recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.  However, this condition is stronger than that necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes.}.
Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
Nonetheless, we have used the tomography protocol in the next section to numerically verify that $2d-1$ measurement axes suffice for all $d\le120$.

@@

A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\v\Pi_V=\set{\Pi_j^V}$ that will be measured in an experiment, where for shorthand we use a combined index $j\simeq\p{\v v_j,\mu_j}$ that specifies both a measurement axis $\v v_j$ and an outcome $\mu_j$, and for clarify we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
The projectors $\v\Pi_V$ can be used to construct the overlap (Gram) matrix
\begin{align}
  G_V = \sum_{i,j=1}^{\abs{V}\times d} \obk{\Pi_i^V|\Pi_j^V} \op{i}{j}.
  \label{eq:overlap_matrix}
\end{align}
Eigenvectors of $G_V$ define linearly independent linear combinations of the projectors in $\v\Pi_V$.
If the projectors in $\v\Pi_V$ span the entire space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
Indexing these eigenvalues $G^V_k$ and the corresponding (normalized) eigenvectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$,
we can construct the orthonormal operators
\begin{align}
  Q^V_k \equiv \sum_j g^V_{kj} \Pi_j^V,
  &&
  g^V_{kj} \equiv \f{x^V_{kj}}{\sqrt{G^V_k}}.
\end{align}
These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
\begin{align}
  \rho = \sum_{k=1}^{d^2} \rho_k^V Q_k^V,
  &&
  \rho_k^V \equiv \bk{Q^V_k}_\rho.
\end{align}
Given empirical estimates $E_j^V$ of the expectation values $\bk{\Pi_j^V}_\rho$, an empirical estimate $\tilde\rho$ of $\rho$ is then
\begin{align}
  \tilde\rho \equiv \sum_k \tilde\rho^V_k Q^V_k,
  \label{eq:reconstructed_state}
\end{align}
where
\begin{align}
  \tilde\rho^V_k \equiv \sum_j g^V_{kj} E_j^V
  \approx \sum_j g^V_{kj} \bk{\Pi_j^V}_\rho
  = \bk{Q^V_k}_\rho
  = \rho^V_k.
\end{align}
The eigenvalues and eigenvectors of the overlap matrix $G_V$ provide strict upper bounds on the distance between the empirical estimate $\tilde\rho$ and the true state $\rho$.
If the estimates $E^V_j$ approximate $\bk{\Pi^V_j}_\rho$ to within an error $\epsilon$, i.e.
\begin{align}
  \abs{E^V_j-\bk{\Pi^V_j}_\rho} \le \epsilon,
\end{align}
then the error in the estimate $\tilde\rho^V_k$ of $\rho^V_k$ is
\begin{align}
  \abs{\tilde\rho^V_k-\rho^V_k}
  &= \abs{\sum_j g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}} \\
  &\le \sum_j \abs{g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}} \\
  &\le \epsilon \norm{\v g^V_k}_1,
\end{align}
where $\norm{\v g}_1=\sum_j\abs{g_j}$ is the 1-norm of the vector $\v g$.
Using the fact that the operators $Q^V_k$ are orthonormal, we can bound the 2-norm (Euclidean) distance $\norm{\tilde\rho-\rho}_2$ between $\tilde\rho$ and $\rho$ as
\begin{align}
  \norm{\tilde\rho-\rho}_2^2
  = \sum_k \abs{\tilde\rho^V_k - \rho^V_k}^2
  \le \epsilon^2 \N_V^2,
  \label{eq:bound}
\end{align}
where
\begin{align}
  \N_V^2 \equiv \min_U \sum_k \norm{U\v g^V_k}_1^2
  = \min_U \sum_k \f{\norm{U\v x^V_k}_1^2}{G^V_k}
  \label{eq:norm}
\end{align}
is a norm uniquely determined by the set of measurement axes $V$, with minimization performed over unitary transformations $U$ of all $\v g^V_k$ and $\v x^V_k$ to account for the fact that their 1-norms are generally basis-dependent.
Seen otherwise, this minimization accounts for the fact that we can repeat the analysis between Eqs.~\eqref{eq:overlap_matrix} and \eqref{eq:bound} with any $G_V\to U G_V U^\dag$, where $U$ is an arbitrary unitary that mixes the projectors in $\v\Pi_V$.
[\red{Actually, I don't think this minimization is quite correct as stated, but it won't matter after some pending changes to this section.  Rather than an absolute error $\epsilon$, we can assume that $E_j$ approximates $\bk{\Pi_j}_\rho$ up to Gaussian noise with variance $\epsilon^2$.  In that case, I think the mean (squared) distance between $\tilde\rho$ and $\rho$ is $\epsilon^2\sum_k1/G^V_k$, which in retrospect is a much nicer result.  I am still working to write this up, which will involve some changes to this section.}]

To better understand the bound in Eq.~\eqref{eq:bound}, we observe that by a change of operator basis, or equivalently a particular choice of unitary $U_T$, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
Orthogonality of transition operators with different degrees $\ell$ then implies that the overlap matrix
\begin{align}
  G_V^{\t{trn}} \equiv U_T G_V U_T^\dag
  = \sum_{\ell,\v v,\v v'} \obk{T_{\v v\ell,0}|T_{\v v'\ell,0}}
  \op{\v v\ell}{\v v'\ell}
\end{align}
has a block-diagonal structure, with $d$ blocks of size $\abs{V}\times\abs{V}$.
Eigenvectors of $G_V$ can therefore have at most $\abs{V}$ nonzero components in the transition operator basis, which implies that
\begin{align}
  \N_V \le \sum_k \f{\norm{U_T\v x^V_k}_1^2}{G^V_k}
  \le \sum_k \f{\abs{V}\,\norm{U_T\v x^V_k}_2^2}{G^V_k}
  = \sum_k \f{\abs{V}}{G^V_k}.
\end{align}
Note that these sums are performed over the $d^2$ largest eigenvalues of $G_V$ (or $G_V^{\t{trn}}$), and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
Altogether, diagonalizing each of $d$ blocks of $G_V^{\t{trn}}$ takes a time that is $O(\abs{V}^3)$, so an upper bound for $\N_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.

The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the cost function
\begin{align}
  \tilde\N_V^2 \equiv \sum_k \f{\norm{\v y^V_k}_1^2}{G^V_k}
  \ge \N_V^2,
\end{align}
where $\v y^V_k$ are the (normalized) eigenvectors of the overlap matrix $G_V^{\t{trn}}$ in the transition operator basis, with corresponding eigenvalues $G^V_k$.
If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\tilde\N_V^2$.
Upon choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi^V_j}_\rho$.
If these measurements can approximate the expectation values $\bk{\Pi^V_j}_\rho$ to within a known error $\epsilon$, then the empirical estimate $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance of $\epsilon\N_V$.
If the empirical estimate $\tilde\rho$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced via maximum-likelihood corrections \cite{smolin2012efficient}.
In terms of classical computing costs before and after estimating the expectation values $\bk{\Pi^V_j}_\rho$, this tomography protocol is bottlenecked by the need to diagonalize the overlap matrix $G_V^{\t{trn}}$, which has $d$ blocks of size $\abs{V}\times\abs{V}$.
Altogether, this protocol therefore has an $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.

@@

As a last point, we mention that the cost function $\tilde\N_V^2$ can also evaluate the quality of measurement axes chosen using existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-cost sets of measurement axes with higher probability.
Two observations may explain the poor performance of the sparse signal recovery methods.
First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely with a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of sparse signals.
Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How many measurement axes?}
\label{sec:axes}
